---
title: "CBS Pest Recruitment Analysis"
author: "Shihong Jia"
date: "Apr 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Analysis 2: Recruitment analysis
The presentation of the relationship between conspecific adult density and seedling survival under three pesticide treatments (i.e. fungicide, insecticide and control). Both prediction and the confidence intervals and the observed values will be shown in one graph.


### Load packages
```{r loadlibs, warning=FALSE, message=FALSE}
library(ggplot2)
library(cowplot)
library(lme4)
library(pbkrtest)
library(tidyr)
library(dplyr)
library(optimx)
library(broom)
library(glmmTMB)
```



### Data and models
```{r data and models}
## read data
pest.recr.dat0 <- read.csv("data/pest.recr.dat.csv")


pest.recr.dat0$pesticide <- factor(pest.recr.dat0$pesticide, levels=c('W', 'F', 'I'))
pest.recr.dat0$census <- factor(pest.recr.dat0$census, levels=c('16sp', '17sp'))
pest.recr.dat0$exclosure <- factor(pest.recr.dat0$exclosure)
pest.recr.dat0$site <- factor(pest.recr.dat0$site)
pest.recr.dat0$quad.unique <- factor(pest.recr.dat0$quad.unique)

pest.recr.dat <- pest.recr.dat0[, -c(1,11:13,15)]

str(pest.recr.dat)



##---------------- model -----------------##

### all seedlings were pooled
## previous analyses had shown that there was no significant interaction effect 
## between exclosure and other variables
mod.recr1 <- glmer(recruits ~ pesticide*(lg_A.con + lg_A.het) + exclosure +  census + site  +
                     (1+census|quad.unique) + (1|sp.),data=pest.recr.dat, family=poisson)


# not consider exclosure in the further analyses
mod.recr2 <- glmer(recruits ~ pesticide*(lg_A.con + lg_A.het) + census + site  +
                     (1+census|quad.unique) + (1|sp.),data=pest.recr.dat, family=poisson)

# do not consider census as the random-slope effect
mod.recr3 <- glmer(recruits ~ pesticide*(lg_A.con + lg_A.het) + census + site  +
                     (1|quad.unique) + (1|sp.),data=pest.recr.dat, family=poisson)

# do not include census and site as fixed effects, and not consider census as a random-slope effect
mod.recr4 <- glmer(recruits ~ pesticide*(lg_A.con + lg_A.het) + 
                     (1|quad.unique) + (1|sp.),data=pest.recr.dat, family=poisson)


anova(mod.recr1, mod.recr2, mod.recr3, mod.recr4)

# present the best model
summary(mod.recr2, correlation=FALSE)

# how about mod.recr4?
summary(mod.recr4, correlation=FALSE)

```


### Prediction1
The experience in the survival analysis told me I should try the mod.recr4 (i.e. without considering census and site as fixed effects and census as a random-slope effect). Also, I would like to predict without considering random effects.
```{r recruitmentprediction}
# First, I need to remove some potential outliers
pest.recr.dat1 <- subset(pest.recr.dat, A.con < 3000)
dim(pest.recr.dat1); dim(pest.recr.dat)


# re-fit the model
mod.recr4_rf <- glmer(recruits ~ pesticide*(lg_A.con + lg_A.het) + 
                     (1|quad.unique) + (1|sp.),data=pest.recr.dat1, family=poisson)

summary(mod.recr4_rf, correlation=FALSE)



## prediction

# temporary data 
recr.tmpdat <- pest.recr.dat1[, c("pesticide", "lg_A.con", "lg_A.het", "quad.unique", "sp.")]
dim(recr.tmpdat)


recr.jvalues <- with(pest.recr.dat1, seq(from = min(lg_A.con), to = max(lg_A.con), length.out = 100))




# First, I will show you the results without considering the random effects
# calculate predicted recruits and store in a list 

recr.biprobs_norand <- lapply(levels(pest.recr.dat1$pesticide), function(treat) 
	{recr.tmpdat$pesticide[] <- treat 
	lapply(recr.jvalues, function(j) {
	recr.tmpdat$lg_A.con <- j 
	predict(mod.recr4_rf, newdata = recr.tmpdat, re.form=~0, type = "response") 
  }) 
})

# get means and quartiles for all jvalues for each level of Pesticide 
recr.plotdat_norand <- lapply(recr.biprobs_norand, function(X) {
	 recr.temp <- t(sapply(X, function(x) {
	 c(M=mean(x), quantile(x, c(.25, .75))) 
	}))
	recr.temp <- as.data.frame(cbind(recr.temp, recr.jvalues))
	colnames(recr.temp) <- c("Predicted", "Lower", "Upper", "Lg_A.con") 
	return(recr.temp)
})


# collapse to one data frame 
recr.plotdat_norand <- do.call(rbind, recr.plotdat_norand) 

# add pesticide treatment
recr.plotdat_norand$pesticide <- factor(rep(levels(pest.recr.dat1$pesticide), 
					each = length(recr.jvalues)))


# back to the original data of x
recr_mean_logx <- mean(log(pest.recr.dat1$A.con + 1))
recr_sd_logx <- sd(log(pest.recr.dat1$A.con + 1))
recr.plotdat_norand$x <- exp((recr.plotdat_norand$Lg_A.con*recr_sd_logx) + recr_mean_logx)-1


# re-level the pesticide treatment
recr.plotdat_norand$pesticide <- factor(recr.plotdat_norand$pesticide, levels=c("W","F","I"))
summary(recr.plotdat_norand)



## ------------------ the observed values -----------------##
## generate k sets of simulated y at these new points.
ob.recr.newdat <- with(pest.recr.dat1, expand.grid(lg_A.con=lg_A.con, pesticide=levels(pesticide), 
		lg_A.het=mean(lg_A.het)))

dim(ob.recr.newdat)


ob.recr.newdat$pred <- predict(mod.recr4_rf, newdata=ob.recr.newdat, re.form=~0, type='link')
pest.recr.dat1$resd <- resid(mod.recr4_rf)


# select the lg_A.con in the original data
ob.recr.newdat$resd <- pest.recr.dat1[pest.recr.dat1$lg_A.con %in% ob.recr.newdat$lg_A.con,]$resd


ob.recr.newdat$obse_lk <- ob.recr.newdat$pred + ob.recr.newdat$resd
ob.recr.newdat$obse <- exp(ob.recr.newdat$obse_lk)
summary(ob.recr.newdat)


# replace the pesticide
ob.recr.newdat$pesticide <- pest.recr.dat1[pest.recr.dat1$lg_A.con %in% ob.recr.newdat$lg_A.con,]$pesticide


# back to the original data of x
ob.recr.newdat$x <- exp((ob.recr.newdat$lg_A.con*recr_sd_logx) + recr_mean_logx)-1
summary(ob.recr.newdat); dim(ob.recr.newdat)



## bin the x to 300 density
ob.recr.newdat$A.con_cat <- cut(ob.recr.newdat$x, seq(-1, 2999, 300))

obse.recr_cat <- summarise(group_by(ob.recr.newdat, A.con_cat, pesticide), 
			mean = mean(obse), se=sd(obse)/sqrt(length(obse)))

# change the A.con_cat to continuous data
obse.recr_cat$x <- rep(seq(150,2850,300), each=3)

summary(obse.recr_cat)


# plot
ggplot() +
  geom_line(data=recr.plotdat_norand, aes(x=x, y=Predicted, colour=pesticide),size=0.8) +
  geom_ribbon(data=recr.plotdat_norand, aes(x=x, ymin = Lower, ymax = Upper, fill = pesticide), 
              alpha = 0.2) +
  geom_point(data=obse.recr_cat,aes(x=x, y = mean, color=pesticide),
		size=2,position=position_dodge(width = 115)) +
  geom_errorbar(data=obse.recr_cat,aes(x=x, ymin=mean-se, ymax=mean+se,color=pesticide),
		width=0.1, position=position_dodge(width = 115)) +
  scale_color_manual(values = c('gray40','orangered','limegreen'),
	labels=c("Control","Fungicide","Insecticide")) + 
  scale_fill_manual(values = c('gray40','orangered','limegreen'),
	labels=c("Control","Fungicide","Insecticide")) +
  labs(x="Conspecific adult density (m2)", y="Recruitment", 
  colour="Pesticide",linetype="Pesticide", fill="Pesticide") +
  theme(legend.title=element_blank(), legend.background = element_rect(fill="lightgray"),
		legend.position ="right") + theme_bw()


```

The prediction are really underestimated compared with the observed values.



## Prediction2
```{r prediction with random effects}

recr.biprobs <- lapply(levels(pest.recr.dat1$pesticide), function(treat) 
	{recr.tmpdat$pesticide[] <- treat 
	lapply(recr.jvalues, function(j) {
	recr.tmpdat$lg_A.con <- j 
	# with random effects
	# predict(mod.recr4_rf, newdata = recr.tmpdat, re.form=~0, type = "response") 
	predict(mod.recr4_rf, newdata = recr.tmpdat, type = "response") 
  }) 
})

# get means and quartiles for all jvalues for each level of Pesticide 
recr.plotdat <- lapply(recr.biprobs, function(X) {
	 recr.temp <- t(sapply(X, function(x) {
	 c(M=mean(x), quantile(x, c(.25, .75))) 
	}))
	recr.temp <- as.data.frame(cbind(recr.temp, recr.jvalues))
	colnames(recr.temp) <- c("Predicted", "Lower", "Upper", "Lg_A.con") 
	return(recr.temp)
})


# collapse to one data frame 
recr.plotdat <- do.call(rbind, recr.plotdat) 

# add pesticide treatment
recr.plotdat$pesticide <- factor(rep(levels(pest.recr.dat1$pesticide), 
					each = length(recr.jvalues)))


# back to the original data of x
recr.plotdat$x <- exp((recr.plotdat$Lg_A.con*recr_sd_logx) + recr_mean_logx)-1


# re-level the pesticide treatment
recr.plotdat$pesticide <- factor(recr.plotdat$pesticide, levels=c("W","F","I"))
summary(recr.plotdat)


# plot with observed values
ggplot() +
  geom_line(data=recr.plotdat, aes(x=x, y=Predicted, colour=pesticide),size=0.8) +
  geom_ribbon(data=recr.plotdat, aes(x=x, ymin = Lower, ymax = Upper, fill = pesticide), 
              alpha = 0.2) +
  geom_point(data=obse.recr_cat,aes(x=x, y = mean, color=pesticide),
		size=2,position=position_dodge(width = 115)) +
  geom_errorbar(data=obse.recr_cat,aes(x=x, ymin=mean-se, ymax=mean+se,color=pesticide),
		width=0.1, position=position_dodge(width = 115)) +
  scale_color_manual(values = c('gray40','orangered','limegreen'),
	labels=c("Control","Fungicide","Insecticide")) + 
  scale_fill_manual(values = c('gray40','orangered','limegreen'),
	labels=c("Control","Fungicide","Insecticide")) +
  labs(x="Conspecific adult density (m2)", y="Recruitment", 
  colour="Pesticide",linetype="Pesticide", fill="Pesticide") +
  theme(legend.title=element_blank(), legend.background = element_rect(fill="lightgray"),
		legend.position ="right") + theme_bw()


```

It is better to predict with random effects.


